---
title: "218X Assignment 2"
author: "Lisiman Hua"
date: "10/4/2021"
output: html_document
---
```{r, include = F}
knitr::opts_chunk$set(warning = F, message = F)
```
```{r}
library(tigris)
library(tidyverse)
library(sf)
library(mapview)
library(leaflet)
library(censusapi)
Sys.setenv(CENSUS_KEY="5042355b73b371569968738cfc4f492be94675c1")
```
2020 dateaset
```{r}
dec_vars_2020 <-
  listCensusMetadata(
    name = "2020/dec/pl",
    type = "variables"
    )

smc_pop_2020 <-
  getCensus(
    name = "dec/pl",
    vintage = 2020,
    region = "block:*", 
    regionin = "state:06+county:081",
    vars = "group(P1)"
  ) %>% 
  transmute(
    GEOID20 = paste0(state,county,tract,block),
    pop_2020 = P1_001N
  )
blocks_2020 <- blocks("CA", "San Mateo", year = 2020, progress_bar = F)
```
2010 dataset
```{r}
dec_vars_2010 <-
  listCensusMetadata(
    name = "2010/dec/pl",
    type = "variables"
  )
smc_pop_2010 <-
  getCensus(
    name = "dec/pl",
    vintage = 2010,
    region = "block:*", 
    regionin = "state:06+county:081",
    vars = "group(P1)"
    ) %>% 
  transmute(
    GEOID10 = paste0(state,county,tract,block),
    pop_2010 = P001001
  )
blocks_2010 <- blocks("CA", "San Mateo", year = 2010, progress_bar = F)
```

```{r}
nfo_boundary <- places("CA", progress_bar = F) %>% 
  filter(NAME == "North Fair Oaks")
nfo_blocks_2010 <- blocks_2010 %>% 
  st_centroid() %>% 
  .[nfo_boundary, ] %>% 
  st_drop_geometry() %>% 
  left_join(blocks_2010 %>% select(GEOID10)) %>% 
  st_as_sf() %>% 
  select(GEOID10) %>% 
  left_join(smc_pop_2010)
mapview(nfo_boundary) + mapview(nfo_blocks_2010, alpha.regions = 0, color = "red", lwd = 2)
```
Find all 2020 blocks that intersect with 2010 blocks and add the population column.
```{r}
nfo_blocks_2020 <- blocks_2020 %>%
  .[nfo_boundary, ] %>% 
  select(GEOID20) %>% 
  left_join(smc_pop_2020)
```
spatial subsetting
```{r}
nfo_blocks_2020_intersect <- nfo_blocks_2020 %>% 
  st_transform(26910) %>% 
  mutate(original_area = st_area(.)) %>% 
  st_intersection(
    nfo_blocks_2010 %>%
      select(GEOID10) %>% 
      st_transform(26910)
  ) %>% 
  mutate(
    leftover_area = st_area(.),
    perc_area = as.numeric(leftover_area / original_area),
    pop_2020 = pop_2020 * perc_area
  )
```
At this point, the dataframe has both GEOID20 and GEOID10 IDs. The GEOID10 IDs tell us which GEOID20 pieces can be pieced together to form the full GEOID10 shape. So we can group and summarize to get the 2020 population counts within the GEOID10 boundaries. We don't need to retain geometries at this step, since we can join this back to the 2010 data.
```{r}
nfo_blocks_2020_reshaped <- nfo_blocks_2020_intersect %>%
  st_drop_geometry() %>% 
  group_by(GEOID10) %>% 
  summarize(
    pop_2020 = sum(pop_2020, na.rm=T) %>% round()
  )
```
compute the absolute difference for persons per acre
```{r}
projection <- "+proj=utm +zone=10 +ellps=GRS80 +datum=NAD83 +units=ft +no_defs"
nfo_pop_2010_2020 <- nfo_blocks_2010 %>% 
  left_join(nfo_blocks_2020_reshaped) %>% 
  mutate(
    absolute_pop = (pop_2020 - pop_2010) / as.numeric(st_area(.)) * 43560
  )
```
The SF Chronicle example has a color palette that goes from red negative values to green negative values, with white in the center. To get our color palette to be symmetrical across 0, we need the minimum and maximum bounds to be equidistant from 0. For our absolute change, let's try stretching one bound to meet the other. Using https://www.nceas.ucsb.edu/sites/default/files/2020-04/colorPaletteCheatsheet.pdf, we can identify "PiYG" as a satisfactory palette for our purposes.

```{r}
max <- max(abs(max(nfo_pop_2010_2020$absolute_pop)), abs(min(nfo_pop_2010_2020$pop_2020)))
absolute_pal <- colorNumeric(
  palette = "PiYG",
  domain = c(-max,max)
)
```
map
```{r}
leaflet(nfo_pop_2010_2020) %>% 
  addProviderTiles(provider = providers$CartoDB.Positron) %>% 
  addPolygons(
    fillColor = ~absolute_pal(nfo_pop_2010_2020$absolute_pop),
    label = ~round(nfo_pop_2010_2020$absolute_pop),
    color = "white",
    opacity = 0.5,
    fillOpacity = 0.75,
    weight = 1
  ) %>% 
  addLegend(
    position = "topright",
    pal = absolute_pal,
    value = c(-max,max),
    title = "Absolute Population Change</br>(persons per acre)"
  )
```
