---
title: "10/12 lecture and textbook chart"
author: "Lisiman Hua"
date: "10/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(censusapi)
Sys.setenv(CENSUS_KEY="5042355b73b371569968738cfc4f492be94675c1")


acs_vars_2019_5yr <-
  listCensusMetadata(
    name = "2019/acs/acs5",
    type = "variables"
  )
```

```{r}
bay_income_white_2019 <- 
  getCensus(
  name = "acs/acs5",
  vintage = 2019,
  region = "county:001,013,041,055,075,081,085,095,097", 
  regionin = "state:06",
  vars = "group(B19001A)" #group()means you want to get all the variables in the table
  ) %>%
  select(
    ends_with("E")&-c(state, NAME)
    ) %>%
  summarize_all(sum) %>%
  pivot_longer(
    everything(),
    names_to = "name",
    values_to = "estimate"
  ) %>%
  left_join(
    acs_vars_2019_5yr %>%
      select(name,label)
  ) %>%
  separate(
    label,
    into = c(NA, NA, "income"),
    sep = "!!"
  ) %>%
  select(-name) %>%
  filter(!is.na(income)) %>%
  mutate(race = "white alone")
```


full_dataset <- NULL

for(x in 1:7){
  # temp <- your code
  full_dataset <-
    rbind(full_dataset, temp)
}


```{r}
race_categories <- c(
  "White alone",
  "Black or African American alone",
  "American Indian and Alaska Native alone",
  "Asian alone",
  "Native Hawaiian and Other Pacific Islander alone",
  "Some Other Race alone",
  "Two or more Races"
)

bay_income_race_2019 <-
  1:7 %>% 
  map_dfr(function(x){
    getCensus(
  name = "acs/acs5",
  vintage = 2019,
  region = "county:001,013,041,055,075,081,085,095,097", 
  regionin = "state:06",
  vars = paste0("group(B19001",LETTERS[x],")") 
  ) %>%
  select(
    ends_with("E")&-c(state, NAME)
    ) %>%
  summarize_all(sum) %>%
  pivot_longer(
    everything(),
    names_to = "name",
    values_to = "estimate"
  ) %>%
  left_join(
    acs_vars_2019_5yr %>%
      select(name,label)
  ) %>%
  separate(
    label,
    into = c(NA, NA, "income"),
    sep = "!!"
  ) %>%
  select(-name) %>%
  filter(!is.na(income))%>%
  mutate(race = race_categories[x])
  })

```
```{r}
race_categories <- c(
  "White alone",
  "Black or African American alone",
  "American Indian and Alaska Native alone",
  "Asian alone",
  "Native Hawaiian and Other Pacific Islander alone",
  "Some Other Race alone",
  "Two or more Races"
)

bay_income_race_2019 <-
  1:7 %>% 
  map_dfr(function(x){
    getCensus(
  name = "acs/acs5",
  vintage = 2019,
  region = "county:001,013,041,055,075,081,085,095,097", 
  regionin = "state:06",
  vars = paste0("group(B19001",LETTERS[x],")") 
  ) %>%
  select(
    ends_with("E")&-c(state, NAME)
    ) %>%
  summarize_all(sum) %>%
  pivot_longer(
    everything(),
    names_to = "name",
    values_to = "estimate"
  ) %>%
  left_join(
    acs_vars_2019_5yr %>%
      select(name,label)
  ) %>%
  separate(
    label,
    into = c(NA, NA, "income"),
    sep = "!!"
  ) %>%
  select(-name) %>%
  filter(!is.na(income))%>%
  mutate(race = race_categories[x])
  })
```


```{r}
bay_income_race_2019 %>% 
  group_by(income, race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = income %>% factor(levels = unique(bay_income_race_2019$income)),
      y = estimate,
      fill = race %>% factor(levels = unique(bay_income_race_2019$race))
    ),
    stat = "identity",
    position = "stack"
  ) +
  labs(
    x = "Household income",
    y = "Number of households",
    title = "Bay Area household income by race",
    fill = "Race of householder"
  ) +
  coord_flip()
```

```{r}
bay_income_race_2019 %>% 
  group_by(income, race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = income %>% factor(levels = rev(unique(bay_income_race_2019$income))),
      y = estimate,
      fill = race %>% factor(levels = rev(unique(bay_income_race_2019$race)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Household income",
    y = "Proportion of households",
    title = "Bay Area household income by race",
    fill = "Race of householder"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )
```
```{r}
bay_race_total_2019 <-
  bay_income_race_2019 %>% 
  group_by(race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  mutate(income = "Total")

bay_income_race_2019 %>% 
  group_by(income, race) %>% 
  summarize(estimate = sum(estimate)) %>% 
  rbind(bay_race_total_2019) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = income %>% factor(levels = rev(c("Total",unique(bay_income_race_2019$income)))),
      y = estimate,
      fill = race %>% factor(levels = rev(unique(bay_income_race_2019$race)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Household income",
    y = "Proportion of households",
    title = "Bay Area household income by race",
    fill = "Race of householder"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )
```



